---
title: "Pareto models for top-incomes"
author: "Arthur Charpentier & Emmanuel Flachaire"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("/Users/acharpen/Dropbox (Télécom SudParis)")
library(devtools)
#devtools::install_github("freakonometrics/TopIncome")
#library(TopIncome)
library(Hmisc)
library(knitr)
latexImg = function(latex){
    link = paste0('http://latex.codecogs.com/gif.latex?',
           gsub('\\=','%3D',URLencode(latex)))
    link = gsub("(%..)","\\U\\1",link,perl=TRUE)
    return(paste0('![](',link,')'))
}
```

# Install the `TopIncome` library

The `TopIncome` library can be installed from *github*,

```{r}
library(devtools)
devtools::install_github("freakonometrics/TopIncomes")
library(TopIncomes)
```

# Fitting Pareto Models

```{r}
n <- 1000
set.seed(123)
x <- repd(n,.5,1,-1)
w <- rgamma(n,10,10)
```

## Pareto 1

The **Pareto type 1** distribution is bounded from below by $u>0$, and with tail parameter `r latexImg('\\alpha')` it has the cumulative distribution function
`r latexImg('\\displaystyle{F(x)=1-\\left(\\frac{x}{u}\\right)^{-\\alpha}}')` for `r latexImg('x\\geq u')`. Note that the tail index is `r latexImg('\\xi=1/\\alpha')`.

```{r}
estim <- MLE.pareto1(data=x, weights=w, threshold=1)
estim
```

## Generalized Pareto 

The **Generalized Pareto** distribution is bounded from below by $u>0$, with tail parameter `r latexImg('\\alpha')` : the cumulative distribution function is
`r latexImg('\\displaystyle{F(x)=1-\\left[1+\\left(\\frac{x-u}{\\sigma}\\right)\\right]^{-\\alpha}}')` for `r latexImg('x\\geq u')`. Note that the tail index is `r latexImg('\\xi=1/\\alpha')`.

```{r}
estim <- MLE.gpd(data=x, weights=w, threshold=1)
estim
```

## Extended Pareto 

The **Extended Pareto** distribution is bounded from below by $u>0$, and has cumulative distribution function
`r latexImg('\\displaystyle{F(x)=1-\\left[\\frac{x}{u}\\left(1+\\delta-\\delta\\left(\\frac{x}{u}\\right)^\\tau\\right)\\right]^{-\\alpha} }')` for `r latexImg('x\\geq u')`. Note that the tail index is `r latexImg('\\xi=1/\\alpha')`.

```{r}
estim <- EPD(data=x, weights=w)
estim
```

# Application to Income

Consider some simulated data,

```{r}
url_1 <- "https://github.com/freakonometrics/TopIncome/raw/master/data_csv/dataframe_yw_1.csv"
df <- read.table(url_1,sep=";",header=TRUE)
data_1  <-  tidy_income(income = df$y, weights = df$w)
Pareto_diagram(data_1)
```

```{r}
T <- Table_Top_Share(data_1, p=.01)
```

Tail index `r latexImg('\\alpha')`, for three fited distributions

```{r}
T$TailIndex
```

```{r, echo=FALSE}
library(knitr)
dt=data.frame(T$TailIndex)[-1,]
q=c(.1,.05,.01)
names(dt)=paste("top",round(100*(1-q)),"%",sep="")
rownames(dt)=c("Pareto_1","GPD","EPD")
kable(dt,caption="Tail Index (alpha)",bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

Top share income, for three fited distributions

```{r}
T$TopShare
```

```{r, echo=FALSE}
dt=data.frame(T$TopShare*100)[-1,]
q=c(.1,.05,.01)
names(dt)=paste("top",round(100*(1-q)),"%",sep="")
rownames(dt)=c("EDF","Pareto_1","GPD","EPD")
kable(dt,caption="Top Share (in percent)", bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

See also (to get automatically tables in a markdown format)

```{r}
# T <- Table_Top_Share(data_1, p=.01, md=TRUE)
```


```{r}
Top_Incomes(data_1)
```
